---
title: "Math 312 Midterm"
author: 'Rishi Gummakonda'
output: pdf_document
---

## Question 1

Code is taken from Homework 3 Question 1,  Lab 4 Question 2, Example 3.2 from Class

```{r}

f_dist <- function (x) {
  return(2*x*exp(-x^2))
}

#integrated
#capital_Fx = -exp(-x^2)

f_inv <- function(x){
  
  return(sqrt(-log(1-x,base=exp(1))))
  #or sqrt(log(1/x))
  
}
 
n <- 1000
u <- runif(n)
f <- f_inv(u)
hist(x=f,col='grey', breaks = 15, border=F, freq=F)
curve(f_dist(x), col='red', add=T)


```


## Question 2
From Homework 3 Question 3.5, and Lab 7 Question 2

```{r}
n <- 1000
u <- runif(n)
x <- numeric(n)
p <- c(0.1,0.3,
       0.4,0.5,
       0.65,0.80,
       0.85,0.95,
       1)
values <- c(0,1,2,3,4,5,6,7,8)


for(i in 1:n)
{
  if(u[i] <=p[1])
  {
    x[i] = 0 
  }
  else if(p[1] <u[i] && u[i] <= p[2])
  {
    x[i] = 2
  }
  
  
  else if(p[2] <u[i] && u[i] <= p[3])
  {
    x[i] = 2
  }
  else if(p[3] <u[i] && u[i] <= p[4])
  {
    x[i] = 3
  }
  else if(p[4] <u[i] && u[i] <= p[5])
  {
    x[i] = 4
  }
  
  else if(p[5] <u[i] && u[i] <= p[6])
  {
    x[i] = 5
  }
  
    else if(p[6] <u[i] && u[i] <= p[7])
  {
    x[i] = 6
    }
    else if(p[7] <u[i] && u[i] <= p[8])
  {
    x[i] = 7
    }
  else if(u[i]>p[8]){
    x[i] =8
  }
  
}
table(x)/n

```
These probabilities seem to match up with the theoretical probabilities pretty well. 


## Question 3
```{r}

func_x <- function(v){
  return(rchisq(v))
}

func_y <- function(v)
{
  return(1/func_x(v))
}

func_y_app <- function(v,df)
{
  #return(dinvchisq())
}

# I know this is somewhere related to Lab 5 and Homework 2but I'm unsure

```

## Question 4
Taken from Homework 4, Question 5, Lab 6, Example 3.7
```{r}
n <- 1000
acceptance_rejection_f <- function(n)
{
  y=numeric(n)
  k=0
  j=0
  while(k<n)
  {
    j=j+1
    x=runif(1,0,pi)
    u=runif(1,0,pi)
    if((12*x^2*sin(x)^2/(2*pi^3-3*pi))>u){
      #we accept x
      k=k+1
      y[k]=x
    }
  }
  return(y)
}
hist(acceptance_rejection_f(n), col='red', breaks=25, freq=F)
curve(
  (12*x^2*sin(x)^2/(2*pi^3-3*pi)),add=T, col="blue"
  )

```

## Question 5
Taken from Homework 5 Question 2, Week 6 Lab 8
```{r}

m = 10000
x<- runif(m, min=0,max=pi)
theta.hat <- mean((x^2+x^3)*(cos(x)^2)
                  *pi)
print(theta.hat)
#print( ((0^2+0^3)*(cos(0)^2)) - ((pi^2+pi^3)*(cos(pi)^2)) )
# I don't know if thats right with the pi

```

## Question 6
Homework 6, Questions 2 and 3, Example Question Lab 8 Antithetic Variate Method

I did part c first and then took averages for theta1 and theta2 to get the answers for a and b
```{r}
n <- 1000
theta1 <-integer(n)
theta2 <- integer(n)
g <- function(x){
   return((x^2+x^3)*exp(x^2-1))
}

for(i in 1:n)
{
  theta1[i] <- mean(g(runif(n)))*1
  u2 <- runif(n/2)
  y1 <- g(u2)
  y2 <- g(1-u2)
  theta2[i] <-mean((y1+y2)/2)
}

# a 
print((mean(g(runif(n)))*1))

#b
print(mean((y1+y2)/2))

#to get a and b i could also get the mean of the 1000 trial runs

#C
#percent reduction in variance
print((var(theta1)-var(theta2))/var(theta1))




```

## Question 7
Week 7 Lab 11 and Homework 6 Question 4
```{r}

# a
m <- 1000
g <- function (x)
  {
  return(exp(-x)/(1+x^2)*I(x>=1)*I(x<=Inf))
}
x <- rexp(m,1)

fg <- g(x)/exp(-x)
theta.hat <- mean(fg)
se <- sd(fg)

print(theta.hat)

#b
#copying from example code from class
x2 <- rcauchy(m)
i <- c(which(x2>1),which(x2<0))
x2[i] <- 2
fg2 <- g(x2)/dcauchy(x2)
theta.hat2 <- mean(fg2)
se2 <- sd(fg)

print(theta.hat2)


#c
theta01 <-integer(n)
theta02<- integer(n)

for(i in 1:n)
{
  x <- rexp(m,1)
  fg <- g(x)/exp(-x)
  theta01[i] <- mean(fg)
  
  x2 <- rcauchy(m)
  i <- c(which(x2>1),which(x2<0))
  x2[i] <- 2
  fg2 <- g(x2)/dcauchy(x2)
  theta02[i] <- mean(fg2)
}

print(var(theta01))
print(var(theta02))


#according to this the Exponential delivered the smallest variance


```

