---
title: "Homework-8"
author: "Noah Backman"
date: "10/31/2021"
output: pdf_document
---


## Question 1: Exercise 7.2
```{r}
n <- 20
m <- 1000
mu0 <- 500
sigma <- 100
mu <- seq(350,650,10) #alternatives
M <- length(mu)
power <- numeric(M)
for( i in 1:M ){
  mu1 <- mu[i]
  pvalues <- replicate(m, expr = {
    #simulate under alternative mu1
    x <- rnorm(n,mean=mu1,sd=sigma)
    ttest <- t.test(x,alternative="two.sided",mu=mu0)
    ttest$p.value
  })
  power[i] <- mean(pvalues <= .05)
}
se <- sqrt(power*(1-power)/m)

# plot the empirical power curve
# adding vertical error bars at pi(theta) +/- 2se(pi(theta))

library(ggplot2)
df <- data.frame(mean=mu, power=power, upper=power+2*se, lower=power-2*se)
ggplot(df,aes(x=mean,y=power)) +
  geom_line()+
  geom_vline(xintercept=500,lty=2) +
  geom_hline(yintercept=c(0,.05),lty=1:2) +
  geom_errorbar(aes(ymin=lower, ymax=upper), width=0.2, lwd=1.5)


```


## Question 2: Exercise 7.3
```{r}
set.seed(42)
n <- seq(10, 50, 10)
m <- 1000
mu0 <- 500
sigma <- 100
power_matrix = matrix(NA, nrow=21, ncol=5)
mu <- seq(450,650,10) #alternatives
M <- length(mu)
power <- numeric(M)
count = 1
for (num in n){
  for( i in 1:M ){
    mu1 <- mu[i]
    pvalues <- replicate(m, expr = {
      #simulate under alternative mu1
      x <- rnorm(num,mean=mu1,sd=sigma)
      ttest <- t.test(x,alternative="greater",mu=mu0)
      ttest$p.value
    })
    power[i] <- mean(pvalues <= .05)
  }
  se <- sqrt(power*(1-power)/m)
  power_matrix[,count] = power
  count = count + 1
}
# plot the empirical power curve
# adding vertical error bars at pi(theta) +/- 2se(pi(theta))

library(ggplot2)
df <- data.frame(mean=mu, p0=power_matrix[,1], p1=power_matrix[,2], p2=power_matrix[,3],
                 p3=power_matrix[,4], p4=power_matrix[,5])
ggplot(df) +
geom_line(aes(x=mean, y=p0), color="red") +
geom_line(aes(x=mean, y=p1), color="blue") +
geom_line(aes(x=mean, y=p2), color="green") +
geom_line(aes(x=mean, y=p3), color="yellow") +
geom_line(aes(x=mean, y=p4), color="orange") +
geom_vline(xintercept=500,lty=2) +
geom_hline(yintercept=c(0,.05),lty=1:2)


```


## Question 3: Exercise 11.1
```{r}
f <- function(x, sigma) {
  if (any(x < 0)) return (0)
    stopifnot(sigma > 0)
  return((x / sigma^2) * exp(-x^2 / (2*sigma^2)))
}  
m <- 10000
sigma <- 2
x <- numeric(m)
x[1] <- rchisq(1, df=1)
k <- 0
u <- runif(m)

for (i in 2:m) {
  xt <- x[i-1]
  y <- rchisq(1, df = xt)
  num <- f(y, sigma) * dchisq(xt, df = y)
  den <- f(xt, sigma) * dchisq(y, df = xt)
  if (u[i] <= num/den) x[i] <- y else {
     x[i] <- xt
     k <- k+1     #y is rejected
   }
}

print(k)

index <- 5000:5500
y1 <- x[index]
plot(index, y1, type="l", main="", ylab="x")

```
Comparing the results when sigma=2 compared to when sigma=4, we see that for this
case, when sigma=2, the plot has many more flat plateaus where x does not change 
over our index steps. This indicates worse performance, because there are more
points in the simulation where there is little variation between steps,
and the candidate is rejected. However, there is slightly lower variance in the 
x range for sigma=2. We can see that the range here is from about [1, 7], an 
in the textbook example, we see the x range is from about [2, 10]. Comparing 
also, the difference between number of rejections, we see there are 4009 from the
textbook, and from our runs, there are around 5200, proving the performance is worse.



## Question 4: Exercise 11.2
```{r}
f <- function(x, sigma) {
  if (any(x < 0)) return (0)
    stopifnot(sigma > 0)
  return((x / sigma^2) * exp(-x^2 / (2*sigma^2)))
}

m <- 10000
sigma <- 4
x <- numeric(m)
x[1] <- rchisq(1, df=1)
k <- 0
u <- runif(m)

for (i in 2:m) {
  xt <- x[i-1]
  y <- rgamma(1, xt)
  num <- f(y, sigma) * dgamma(xt, y)
  den <- f(xt, sigma) * dgamma(y, xt)
  if (u[i] <= num/den){
    x[i] <- y 
  }
  else {
    x[i] <- xt
    k <- k+1     #y is rejected
  }
}

print(k)

index <- 5000:5500
y1 <- x[index]
plot(index, y1, type="l", main="", ylab="x")

```
