---
title: "Lab20"
author: "Rishi Gummakonda"
date: "11/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

The scor(bootstrap) data has exam scores in five subjects: mechanics, vectors, algebra, analysis, and
statistics, for 88 students. Compare the five variables in a scatterplot matrix.


```{r }
## Question 1
library(bootstrap)
data(scor)
#pairs(scor[101:150,1:4])

# panel.d plots the densities
panel.d <- function(x, ...){
  usr <- par("usr")
  on.exit(par(usr))
  par(usr=c(usr[1:2], 0, .5))
  lines(density(x))
}

x <- scale(scor[1:88, 1:5])
r <- range(x)
pairs(x, diag.panel=panel.d, xlim= r,ylim= r)

# The lattice package
library(lattice)
splom(scor[1:88, 1:5])

#for all 3 at once, in color, plot 2
scor_groups <- c(scor$mec, scor$vec, scor$alg, scor$ana, scor$sta)
splom(scor[,1:5], groups=scor_groups)

#for all 3 at once, black and white, plot 3
splom(~scor[1:5], groups=scor_groups, data=scor, col=1, pch=c(1,2,3), cex=c(.5, .5, .5))
```

## Question 2
Compute the correlation matrix for the five variables in the scor(bootstrap) data and then use corrplot
to generate a plot to visualize this matrix.

```{r}
#-------------------------#
#--- Correlation Plots ---#
#-------------------------#


## Example 5.2
library(FactoMineR) #decathlon data
library(corrplot)
data("scor")
str(scor) #check the structure of the data


corrMat <- cor(scor) #calculate correlation matrix

#help(corrplot)

corrplot(corrMat, type="upper", tl.col="black", tl.srt=45)

corrplot(corrMat, type="upper", method="square", addCoef.col = "black", diag=F)


```



## Question 3
Use persp function to plot the surface of the following bivariate function

```{r}

## Example 5.3
# the standard BVN density
demo(persp)

##help(outer)
##help(persp)

f <- function(x,y){
  z <- 0.5 * (x^2 +y^3)
}

y <- x <- seq(-3,3, length=50)
z <- outer(x,y,f) # compute density for all (x,y)

persp(x,y,z) # the default plot

persp(x,y,z, theta=45, phi=30, expand=0.6,
      ltheta=120, shade=0.75, ticktype = "detailed",
      xlab="X", ylab="Y", zlab="f(x,y)")

```

