---
title: "Project1"
author: "Matthew Piriya, Rishi Gummakonda, Justin Angelel"
date: "10/31/2021"
output: pdf_document
---
#T-Test Summary:

#Simulation Settings and Code Organization:

Each of our codes were based off of example 7.7 from the reference R code titled
“R Code for Monte Carlo Method in Inference”. We set our alpha or test level to 
0.5. We also set our sample size to 20 and our number of simulations to 1000. 
This way all three simulations we did were conducted under the same 
circumstances and all we were truly comparing was the different distributions.

The code is structured so that variables alpha, n and m are set before the loop
that does the simulation. An array p is also declared to store the p value for 
each iteration of the simulation. Then the loop iterates m times. Within each 
iteration of the loop, the sample of the distribution ($\chi^2$, Uniform or 
exponential) is taken and the two sided t test is performed. Then the p value
is stored in the array p. Following completion of the loop, the p.hat value
is computed along with the squared error for the simulation.

#Findings and Analysis:
 
For the Chi-Squared distribution, we set the mean to 1 because the expected mean
is equal to the degrees of freedom, which is also 1. Then by running 1000 Monte
Carlo simulations of taking 20 random samples from the Chi Squared Distribution,
we found the empirical Type I Error Rate to be 0.09, with a standard error of 
0.009. The Type I Error Rate differed from 0.05 by more than one standard error,
and thus, we can say that the t-test is not robust enough to accommodate for the
departure from normality, or in other words, the $\chi^2$ distribution cannot be
used in place of a normal distribution when taking random samples.

For the Uniform distribution, the process was largely similar, but we instead 
used the Uniform(0, 2) distribution, which also has a mean of 1. Similarly, we
ran 1000 Monte Carlo Simulations of 20 random samples from the uniform distribution,
and found the empirical Type I Error Rate to be 0.045, with a standard error of
0.0066. This time, the empirical Type I Error Rate differed from 0.05 by less
than one standard error, which indicates that the t-test is robust enough to 
the departure from normality between a uniform distribution and a normal one.

For the Exponential distribution, the true mean is still 1, and through the same
process, we found the empirical Type I Error Rate to be 0.075, with a standard
error of 0.0083. Because the Type I Error Rate differed from 0.05 by more than 
one standard error, similar to the $\chi^2$ distribution, we can say that the 
t-test is not robust enough to accommodate for the departure from normality to
the exponential distribution. 

#Appendix with R code:

# i.
```{r}
set.seed(1234)

# based on example 7.7
# number of monte carlo simulations
m <- 1000
# make a vector of length m
p <- numeric(m)
# expected alpha value
alpha <- 0.05

for (i in 1:m){
  # sample from chi square distribution
  u <- rchisq(n=20, df=1)
  # u <- rexp(20,1)
  
  # run t test on u, compare to desired mu = 1
  ttest <- t.test(u, mu=1, alternative="two.sided")

  # store p value
  p[i] <- ttest$p.value
}
# p(Type 1 error) at alpha = .05
p.hat <- mean(p < alpha)
se.hat <- sqrt(p.hat*(1-p.hat)/m)
print(c(p.hat, se.hat))
```

## ii.
```{r}
#Code modified from example 7.7
set.seed(1234) #set seed so results are repeatable
n <- 20 # sample size
alpha <- 0.05 
mu0 <- 1 #compare to mean of 10,000 runiform instances

m <- 1000  #number of replicates
p <- numeric(m)  #storage for p-value
for (j in 1:m){ #run m simulations
  x <- runif(n, 0, 2) #create sample
  ttest <- t.test(x, alternative = "two.sided", mu=mu0) #run t-test
  p[j] <- ttest$p.value #store p value
}

p.hat <- mean(p < alpha)
se.hat <- sqrt(p.hat*(1-p.hat)/m)
print(c(p.hat, se.hat))
```

## iii.
```{r}
## Example 7.7 (Empirical Type I Error Rate)
#Exponential
set.seed(1234)
n <- 20
alpha <- 0.05
mu0 <- 1

m <- 1000  #number of replicates
p <- numeric(m)  #storage for p-value
for (j in 1:m){
  x <- rexp(n, mu0)
  ttest <- t.test(x, alternative = "two.sided", mu=mu0)
  #use two sided because we 
  p[j] <- ttest$p.value
}

p.hat <- mean(p < alpha)
se.hat <- sqrt(p.hat*(1-p.hat)/m)
print(c(p.hat, se.hat))
```

